Probabilistic Symbolic Execution

### 2021-06-25 shared/Paper-Draft/PSE-POPL-2021.tex
WIP : Discussions

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
%%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
\documentclass[acmsmall]{acmart}\settopmatter{}

@@

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
%% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
%% http://ctan.org/pkg/subcaption
\usepackage{mathtools}
\usepackage{algpseudocode,algorithm,algorithmicx}
%\usepackage{amsmath, amssymb,amsthm, stmaryrd}

@@

% Switch Statement Commands
% New definitions
\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicassert{\texttt{assert}}
\algnewcommand\Assert[1]{\State \algorithmicassert(#1)}%
% New "environments"
\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}%
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}%
\algtext*{EndSwitch}%
\algtext*{EndCase}%

@@

@@

	\section{Motivating Example}
	\label{sec:example}

@@

	\subsection{Adding Probabilistic Sampling}
	We now consider the problem of performing symbolic execution on a simple imperative probabilistic programming language, \textbf{pWhile}:
	%
	\[
	P := \mathbf{skip} \mid \mathtt{x} \leftarrow e \mid \mathtt{x} \xleftarrow{\$} d \mid P ; P \mid \mathbf{if}~e~\mathbf{then}~P~\mathbf{else}~P \mid \mathbf{while}~e~\mathbf{do}~P 
	\]

	Above, $\mathtt{x}$ is a program variable, $e$ is an expression, and $d$ is a \textit{discrete} distribution expression. 
	%
	In order to support sampling instructions we make the following additions to traditional symbolic execution:
	%
	\begin{itemize}
		\item \textit{Probabilistic symbolic variables.} We distiguish two types of symbolic variables: \textit{universally quantified} symbolic variables (identical to those in traditional symbolic execution), and \textit{probabilistic} symbolic variables.
		% 
		For each unique sampling instruction a new probabilistic symbolic variable is created to denote the result of sample.
		\item \textit{Distribution map.} We add a new mapping from probabilistic symbolic variables to distribution expressions, $P$, which tracks the distribution from which a probabilistic symbolic variable was originally sampled from.
		\item \textit{Path probability.} For each path, we adjoin a path probability expression, $p$, which is parameterized by universally quantified symbolic variables.
	\end{itemize}

	\begin{algorithm}[H]
		\caption{PSE Assignment Algorithm}
		\begin{algorithmic}[1]
			\Function{PSEAssignment}{$\mathtt{x}, e, \varphi, \sigma, P, I$}
			\Let{$e_{sym}$}{Convert $v$ into a symbolic expression using $\sigma$}
			\State{$\sigma[\mathtt{x}] = e_{sym}$}
			\State\Return{$(\varphi,\sigma, P, I)$}
			\EndFunction
		\end{algorithmic}
	\end{algorithm}

@@

	\begin{algorithm}
		\caption{PSE Branch Algorithm}
		\label{alg:branch}
		\begin{algorithmic}[1]
			\Function{PSEBranch}{$c, \varphi, \sigma, P$}
			\Let{$c_{sym}$}{$\sigma[c]$}
			\Let{$(\delta_1,\ldots,\delta_n)$}{$\dom{P}$}
			\Let{$(d_1,\ldots,d_n)$}{$(P[\delta_1],\ldots,P[\delta_n])$}
			\Let{$p_c$}{$\frac{\displaystyle\sum_{(v_1,\ldots,v_n) \in \dom{d_1} \times \cdots\times \dom{d_n}} [(c_{sym} \wedge \varphi)\{\delta_1 \mapsto v_1,\ldots,\delta_n \mapsto v_n\}]}{\displaystyle\sum_{(v_1,\ldots,v_n) \in \dom{d_1} \times \cdots\times \dom{d_n}}[\varphi\{\delta_1 \mapsto v_1,\ldots,\delta_n \mapsto v_n\}]}$}\\
			\Return{$((\varphi \wedge c_{sym}, \sigma,P,p_c), (\varphi \wedge \neg c_{sym}, \sigma,P,1-p_c))$}
			\EndFunction
		\end{algorithmic}
	\end{algorithm}

	\begin{algorithm}
		\caption{Getting Path Constraints : Symbolic Execution}
		\label{alg:symbex}
		\begin{algorithmic}[1]
			\Function{SymbEx}{$P_{prog} : Program$}
				\Let{$\varphi_{paths}$}{[]}, {$Ex_{stack}$} $\leftarrow$ {[]}, $\Delta \leftarrow$ {[]}, $\phi_{paths} \leftarrow \phi$  \algorithmiccomment{Initialization}
				\Let{$I_{0}$}{\textsc{getStartInstruction}($P_{prog}$)}
				\Let{$S_{0}$}{[$I_{0}$, $\phi_{path}$, $\Delta$]} \algorithmiccomment{Empty Initial State}
				\State {$Ex_{stack}$.\textsc{Push}({$S_{0}$})} \algorithmiccomment{Start with $S_{cur}$ in Execution Stack}
				\While{$Ex_{stack}$ != $\phi$}
					\Let{$S_{cur}$}{$Ex_{stack}$.\textsc{Pop}( )}, $I_{cur} \leftarrow S_{cur}[1]$ \algorithmiccomment{Start State, [$I_{0}$, $\phi_{path}$, $\Delta$]}
					\Switch{\textsc{instType}($I_{cur}$)} \label{line:inst_case}
					\Case{v := e} \algorithmiccomment{Assignment Instruction} \label{line:assign}
						\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}
						\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path}$, $\Delta$[v $\rightarrow$ \textsc{Eval($e$, $\Delta$)}]]}
						\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
					\EndCase
					\Case{\textbf{if} $(c_{sym})$ \textbf{then} $P_{1}$ \textbf{else} $P_{2}$} \algorithmiccomment{Branch Instruction} \label{line:branch}
						\If{(\textsc{isSat}($c_{sym}$) $\wedge$ \textsc{isSat}($\neg c_{sym}$))} \algorithmiccomment{Both cases SAT}
							\Let{$I_{1}$}{\textsc{getStartInstruction}($P_{1}$)} \Let{$I_{2}$} {\textsc{getStartInstruction}($P_{2}$)}	
							\Let{$S_{true}$}{[$I_{1}$, $\phi_{path} \wedge c_{sym}$, $\Delta$]} \label{line:true_state}
							\Let{$S_{false}$}{[$I_{2}$, $\phi_{path} \wedge (\neg c_{sym})$, $\Delta$]} \label{line:false_state}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{false}$})}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{true}$})} \algorithmiccomment{Start with True State}
						\ElsIf{\textsc{isSat}($c_{sym}$)} \algorithmiccomment{True case SAT}
							\Let{$I_{cur}$}{\textsc{getStartInstruction}($P_{1}$)}
							\Let {$S_{cur}$}{[$I_{cur}$, $\phi_{path} \wedge c_{sym}$, $\Delta$]}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
						\ElsIf{\textsc{isSat}($\neg c_{sym}$)} \algorithmiccomment{False case SAT}
							\Let{$I_{cur}$}{\textsc{getStartInstruction}($P_{2}$)}
							\Let{$S_{cur}$}{[$I_{2}$, $\phi_{path} \wedge (\neg c_{sym})$, $\Delta$]}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
						\Else \algorithmiccomment{Unconditional Branch}
							\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}
							\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path}$, $\Delta$]}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
						\EndIf
					\EndCase
					\Case{\textbf{assume}$(v == expr)$} \algorithmiccomment{Assume Instruction} \label{line:assume}
						\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}	
						\If{\textsc{isSat}($v == expr$)} \algorithmiccomment{Check Satisfiablilty}
							\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path} \wedge (v == expr)$, $\Delta$]}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
						\Else 
							%\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path} \wedge $(\textsc{false}), $\Delta$]}
							\Let {$\phi_{path}$}{$\phi$} \algorithmiccomment{Infeasible Path}
						\EndIf
					\EndCase
					\Case{\textbf{HALT}} \algorithmiccomment{Terminate Instruction}
						\Let{$\varphi_{paths}$}{$\varphi_{paths}$.\textsc{Append}($\phi_{path}$)}		
					\EndCase
					\EndSwitch
				\EndWhile \\
				\Return{$\varphi_{paths}$}
			\EndFunction
		\end{algorithmic}
	\end{algorithm}

	When a sampling statement, $\mathtt{x} \xleftarrow{\$} d$, is reached, a fresh probabilistic symbolic variable, $\delta$, is created, $\sigma$ is updated to be $\sigma[\mathtt{x}] = \delta$, and the original distribution $d$ is recorded in $P$ by setting $P[\delta] = d$.
	%
	Note that with the inclusion of probabilistic symbolic variables we can now either branch on universally quantified or probabilistic symbolic variables (or both).
	%
	For guards whose symbolic version does not contain any probabilistic symbolic variables, probabilstic symbolic execution works nearly identically to traditional symbolic execution, save one detail: the probability of taking the branch. 
	%
	If $c_{sym}$ is a symbolic expression which represents the guard to a branch, and $c_{sym}$ does not contain any probabilistic symbolic variables, then we define the probability of taking the ``true'' branch is $[c_{sym}]$, and the probability of taking the ``false'' branch is $[\neg c_{sym}]$, where $[\cdot]$ are Iverson brackets.

	For probabilistic branches, i.e. guards which branch on probabilistic symbolic variables, Alg.~\ref{alg:branch} is used instead of the traditional symbolic execution branch algorithm.
	%
	Without loss of generality, consider a branch of the form $\mathbf{if}~c~\mathbf{then}~P_1~\mathbf{else}~P_2$ where $c$ is a probabilistic branch.
	%
	As before, we define $c_{sym} = \sigma[c]$.
	%
	The core of Alg.~\ref{alg:branch} is computing the probability of taking each branch.
	%
	We interpret probabilistic branches as a conditioning operation on the distributions which are mentioned in the guard of the branch.
	%
	Under this interpretation, we aim to compute $\Pr[c_{sym} \mid \varphi]$ where $\varphi$ is the current path condition formula.

	% Add small example 
	We use a form of model counting in order to compute $p_c$, the probability of $c$ being true.
	%
	Note that each probabilistic symbolic variable, $\delta$, is mapped to exactly one distribution, $d$, and therefore, $\delta \in \dom{d}$.
	%
	So, assuming there are $n$ probabilistic symbolic variables, $\delta_1,\ldots,\delta_n$, and so $n$ distributions, $d_1,\ldots,d_n$, the set of all possible values $\delta_1,\ldots,\delta_n$ be is $\mathcal{D} = \dom{d_1} \times \cdots \times \dom{d_n}$.
	%
	We then count the number of \textit{assignments} from $\mathcal{D}$ which satisfy $c_{sym} \wedge \varphi$ and $\varphi$, and divide these two quantities as shown on line 5 of Alg.~\ref{alg:branch}.
	%
	In Section~\ref{sec:formalization} we have a proof of correctness for this quantity, but the intuition is by applying the definition of conditional probability.
	%
	Note that $p_c$ is not necessarily a value, but rather a symbolic expression containing constants and universally quantified symbolic variables.
	%
	We exploit the fact that the sum of the conditional probabilities of the branch outcomes is 1, which allows us to avoid computing the probability of taking the ``false'' branch directly.

>>>>

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
%%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
\documentclass[acmsmall,review]{acmart}\settopmatter{}

@@

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
%% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
%% http://ctan.org/pkg/subcaption
\usepackage{mathtools}
\usepackage{algpseudocode,algorithm,algorithmicx}
\usepackage{forest}
\usepackage{csquotes}
\usepackage{minted}
\usepackage{multicol}

@@

% For the Case statements
% Switch Statement Commands
% New definitions
\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicassert{\texttt{assert}}
\algnewcommand\Assert[1]{\State \algorithmicassert(#1)}%
% New "environments"
\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}%
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}%
\algtext*{EndSwitch}%
\algtext*{EndCase}%

@@

	One of the goals of symbolic execution is to identify whether a ``bad'' state (e.g. an \texttt{assert} failure) is reached, signifying a bug in the program.
	%
	In probabilistic programs, we are often more interested in either whether a ``bad'' state is reached too often, or, conversely, whether the ``good'' states are reached often enough.

@@

	\section{Motivating Example}
	\label{sec:motiv_example}

	The Monty Hall problem~\cite{selvin1975} is a classic probability puzzle based on the American television show, \textit{Let's Make a Deal}, which showcases how subtle probabilistic reasoning can be.
	%
	The problem itself is simple:
	\begin{displayquote}
		You are a contestant on a gameshow and behind one of three doors there is a car and behind the others, goats.
		% 
		You pick a door and the host, who knows what is behind each of the doors, opens a different door, which has a goat.
		% 
		The host then offers you the choice to switch to the remaining door.
		% 
		Should you?
	\end{displayquote}
	While it may seem unintuitive, regardless of the contestant's original door choice, the contestant who always switches doors will win the car $\frac{2}{3}$ of the time, as opposed to a $\frac{1}{3}$ chance.

	\begin{figure}\centering
		\begin{multicols}{2}
			\inputminted[fontsize=\footnotesize,linenos]{c}{montyhall.c}
		\end{multicols}
		\caption{C code for the Monty Hall Problem}
		\label{fig:montyhall}
	\end{figure}

	We can represent this problem as a probabilistic program, as shown in Fig.~\ref{fig:montyhall}, where $\mathtt{choice} \in [1,3]$ and is the door which is originally chosen by the contestant and $\mathtt{door\_switch}$ is \texttt{true} if the contestant wants to switch doors when asked, and \texttt{false} otherwise.
	%
	Supposedly if $\mathtt{door\_switch} = \mathtt{true}$, regardless of the value of \texttt{choice}, \texttt{monty\_hall} should return \texttt{true} (i.e. the car is won) $\frac{2}{3}$ of the time.
	%
	The problem thus becomes, given a probabilistic program with discrete sampling statements, and a program property, how do we verify that the program satisfies the property?

	One typical solution would be to run a simulation on random inputs and compute the probability of winning if $\mathtt{door\_switch} = \mathtt{true}$, and the probability of winning if $\mathtt{door\_switch} = \mathtt{false}$.
	%
	While for this program this would work as the input space is tiny, the resulting output distributions would be absent any formal guarantees and only be an approximation.

	% 1) What is the problem? Given a program, and a property we wish to check, how do we prove that property?
	% - Introduce the program and the property
	% 2) Given a program, how do we generate the tree with probabilities?
	% 3) Given a tree, how do we make queries against the tree to actually verify the property?

@@

	\subsection{Adding Probabilistic Sampling}
	We now consider the problem of performing symbolic execution on a simple imperative probabilistic programming language, \textbf{pWhile}:
	%
	\[
	S := \mathbf{skip} \mid \mathtt{x} \leftarrow e \mid \mathtt{x} \xleftarrow{\$} d \mid S_1 ; S_2 \mid \mathbf{if}~e~\mathbf{then}~S_1~\mathbf{else}~S_2 \mid \mathbf{while}~e~\mathbf{do}~S 
	\]

	Above, $\mathtt{x}$ is a program variable, $e$ is an expression, and $d$ is a \textit{discrete} distribution expression. 
	%
	In order to support sampling instructions we make the following additions to traditional symbolic execution:
	%
	\begin{itemize}
		\item \textit{Probabilistic symbolic variables.} We distiguish two types of symbolic variables: \textit{universal} symbolic variables (identical to those in traditional symbolic execution), and \textit{probabilistic} symbolic variables.
		% 
		For each unique sampling instruction a new probabilistic symbolic variable is created to denote the result of sample.
		\item \textit{Distribution map.} We add a new mapping from probabilistic symbolic variables to distribution expressions, $P$, which tracks the distribution from which a probabilistic symbolic variable was originally sampled from.
		\item \textit{Path probability.} For each path, we adjoin a path probability expression, $p$, which is parameterized by universal symbolic variables.
	\end{itemize}

	\begin{algorithm}[H]
		\caption{PSE Assignment Algorithm}
		\label{alg:assign}
		\begin{algorithmic}[1]
			\Function{PSEAssignment}{$\mathtt{x}, e, \varphi, \sigma, P$}
			\Let{$e_{sym}$}{$\sigma[e]$}
			\State{$\sigma[\mathtt{x}] = e_{sym}$}
			\State\Return{$(\varphi,\sigma,P)$}
			\EndFunction
		\end{algorithmic}
	\end{algorithm}

	\begin{paragraph}{Assignment}
		For assignment statements of the form $\mathtt{x} \leftarrow e$, where $\mathtt{x}$ is a \textit{program} variable and $e$ is an expression, probabilistic symbolic execution proceeds identically to traditional symbolic execution, as detailed in Alg.~\ref{alg:assign}.
	\end{paragraph}

@@

	\begin{paragraph}{Sampling}
		For sampling statements, $\mathtt{x} \xleftarrow{\$} d$, where $\mathtt{x}$ is a \textit{program} variable and $d$ is a distribution expression, Alg.~\ref{alg:sample} is used.
		% 
		A fresh probabilistic symbolic variable, $\delta$, is created, $\sigma$ is updated to be $\sigma[\mathtt{x}] = \delta$, and the original distribution $d$ is recorded in $P$ by setting $P[\delta] = d$.
	\end{paragraph}

	\begin{algorithm}
		\caption{PSE Branch Algorithm}
		\label{alg:branch}
		\begin{algorithmic}[1]
			\Function{PSEBranch}{$c, \varphi, \sigma, P$}
			\Let{$c_{sym}$}{$\sigma[c]$}
			\Let{$(\delta_1,\ldots,\delta_n)$}{$\dom{P}$}
			\Let{$(d_1,\ldots,d_n)$}{$(P[\delta_1],\ldots,P[\delta_n])$}
			\Let{$p_c$}{$\frac{\displaystyle\sum_{(v_1,\ldots,v_n) \in \dom{d_1} \times \cdots\times \dom{d_n}} [(c_{sym} \wedge \varphi)\{\delta_1 \mapsto v_1,\ldots,\delta_n \mapsto v_n\}]}{\displaystyle\sum_{(v_1,\ldots,v_n) \in \dom{d_1} \times \cdots\times \dom{d_n}}[\varphi\{\delta_1 \mapsto v_1,\ldots,\delta_n \mapsto v_n\}]}$}
			\State\Return{$((\varphi \wedge c_{sym}, \sigma,P,p_c), (\varphi \wedge \neg c_{sym}, \sigma,P,1-p_c))$}
			\EndFunction
		\end{algorithmic}
	\end{algorithm}

	% Symbolic Execution Algorithm.
	\begin{algorithm}
		\caption{Getting Path Constraints : Symbolic Execution}
		\label{alg:symbex}
		\begin{algorithmic}[1]
			\Function{SymbEx}{$P_{prog} : Program$}
			\Let{$\varphi_{paths}$}{[]}, {$Ex_{stack}$} $\leftarrow$ {[]}, $\Delta \leftarrow$ {[]}, $\phi_{paths} \leftarrow \phi$  \algorithmiccomment{Initialization}
			\Let{$I_{0}$}{\textsc{getStartInstruction}($P_{prog}$)}
			\Let{$S_{0}$}{[$I_{0}$, $\phi_{path}$, $\Delta$]} \algorithmiccomment{Empty Initial State}
			\State {$Ex_{stack}$.\textsc{Push}({$S_{0}$})} \algorithmiccomment{Start with $S_{cur}$ in Execution Stack}
			\While{$Ex_{stack}$ != $\phi$}
			\Let{$S_{cur}$}{$Ex_{stack}$.\textsc{Pop}( )}, $I_{cur} \leftarrow S_{cur}[1]$ \algorithmiccomment{Start State, [$I_{0}$, $\phi_{path}$, $\Delta$]}
			\Switch{\textsc{instType}($I_{cur}$)} \label{line:inst_case}
			\Case{v := e} \algorithmiccomment{Assignment Instruction} \label{line:assign}
			\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}
			\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path}$, $\Delta$[v $\rightarrow$ \textsc{Eval($e$, $\Delta$)}]]}
			\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
			\EndCase
			\Case{\textbf{if} $(c_{sym})$ \textbf{then} $P_{1}$ \textbf{else} $P_{2}$} \algorithmiccomment{Branch Instruction} \label{line:branch}
			\If{(\textsc{isSat}($c_{sym}$) $\wedge$ \textsc{isSat}($\neg c_{sym}$))} \algorithmiccomment{Both cases SAT}
			\Let{$I_{1}$}{\textsc{getStartInstruction}($P_{1}$)} \Let{$I_{2}$} {\textsc{getStartInstruction}($P_{2}$)}	
			\Let{$S_{true}$}{[$I_{1}$, $\phi_{path} \wedge c_{sym}$, $\Delta$]} \label{line:true_state}
			\Let{$S_{false}$}{[$I_{2}$, $\phi_{path} \wedge (\neg c_{sym})$, $\Delta$]} \label{line:false_state}
			\State {$Ex_{stack}$.\textsc{Push}({$S_{false}$})}
			\State {$Ex_{stack}$.\textsc{Push}({$S_{true}$})} \algorithmiccomment{Start with True State}
			\ElsIf{\textsc{isSat}($c_{sym}$)} \algorithmiccomment{True case SAT}
			\Let{$I_{cur}$}{\textsc{getStartInstruction}($P_{1}$)}
			\Let {$S_{cur}$}{[$I_{cur}$, $\phi_{path} \wedge c_{sym}$, $\Delta$]}
			\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
			\ElsIf{\textsc{isSat}($\neg c_{sym}$)} \algorithmiccomment{False case SAT}
			\Let{$I_{cur}$}{\textsc{getStartInstruction}($P_{2}$)}
			\Let{$S_{cur}$}{[$I_{2}$, $\phi_{path} \wedge (\neg c_{sym})$, $\Delta$]}
			\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
			\Else \algorithmiccomment{Unconditional Branch}
			\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}
			\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path}$, $\Delta$]}
			\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
			\EndIf
			\EndCase
			\Case{\textbf{HALT}} \algorithmiccomment{Terminate Instruction}
			\Let{$\varphi_{paths}$}{$\varphi_{paths}$.\textsc{Append}($\phi_{path}$)}		
			\EndCase
			\EndSwitch
			\EndWhile \\
			\Return{$\varphi_{paths}$}
			\EndFunction
		\end{algorithmic}
	\end{algorithm}

	\begin{paragraph}{Branches}
		Note that with the inclusion of probabilistic symbolic variables we can now either branch on universal or probabilistic symbolic variables (or both).
		% 
		For guards whose symbolic version does not contain any probabilistic symbolic variables, probabilstic symbolic execution works nearly identically to traditional symbolic execution, save one detail: the probability of taking the branch. 
		% 
		If $c_{sym}$ is a symbolic expression which represents the guard to a branch, and $c_{sym}$ does not contain any probabilistic symbolic variables, then we define the probability of taking the ``true'' branch is $[c_{sym}]$, and the probability of taking the ``false'' branch is $[\neg c_{sym}]$, where $[Q] = 1$ if $Q$ is true, and 0 otherwise (known as Iverson brackets).

		For probabilistic branches, i.e. guards which branch on probabilistic symbolic variables, Alg.~\ref{alg:branch} is used instead of the traditional symbolic execution branch algorithm.
		% 
		Without loss of generality, consider a branch of the form $\mathbf{if}~c~\mathbf{then}~S_1~\mathbf{else}~S_2$ where $c$ is a probabilistic branch.
		% 
		As before, we define $c_{sym} = \sigma[c]$.
		% 
		The core of Alg.~\ref{alg:branch} is computing the probability of taking each branch.
		% 
		We interpret probabilistic branches as a conditioning operation on the distributions which are mentioned in the guard of the branch.
		% 
		Under this interpretation, we aim to compute the conditional probability $\Pr[c_{sym} \mid \varphi] = \frac{\Pr[c_{sym}\wedge \varphi]}{\Pr[\varphi]}$, where $\varphi$ is the current path condition formula.

		We use a form of model counting in order to compute $p_c$, the probability of $c$ being true.
		% 
		Note that each probabilistic symbolic variable, $\delta$, is mapped to exactly one distribution, $d$, and therefore, $\delta \in \dom{d}$.
		% 
		So, assuming there are $n$ probabilistic symbolic variables, $\delta_1,\ldots,\delta_n$, and so $n$ distributions, $d_1,\ldots,d_n$, the set of all possible values $\delta_1,\ldots,\delta_n$ be is $\mathcal{D} = \dom{d_1} \times \cdots \times \dom{d_n}$.
		% 
		We then count the number of elements (or \textit{assignments}) from $\mathcal{D}$ which satisfy $c_{sym} \wedge \varphi$ and $\varphi$, and divide these two quantities as shown on line 5 of Alg.~\ref{alg:branch}.
		% 
		Note that $p_c$ is not necessarily a value, but rather a symbolic expression containing constants and universal symbolic variables.
		% 
		Additionally, we exploit the fact that the sum of the conditional probabilities of the branch outcomes is 1, which allows us to avoid computing the probability of taking the ``false'' branch directly.
	\end{paragraph}

	\begin{figure}[h]
		\centering
		\begin{subfigure}{.3\textwidth}
			\centering
			{\small
				\begin{algorithmic}[1]
					\State{$\mathtt{x}\xleftarrow{\$}\mathrm{UniformInt}(1,3)$}
					\State{$\mathtt{y}\xleftarrow{\$}\mathrm{UniformInt}(1,3)$}
					\If{$\mathtt{x} > 1$}
					\If{$\mathtt{x} < \mathtt{y}$}
					\State\Return{$\mathtt{True}$}
					\EndIf
					\Else
					\State\Return{$\mathtt{False}$}
					\EndIf
				\end{algorithmic}
			}
			\caption{Program}
			\label{fig:exam_program}
		\end{subfigure}
		\begin{subfigure}{.69\textwidth}
			\centering
			{\small
				\begin{forest}baseline,for tree=draw,
					[{$\varphi=\top$},align=center, base=bottom
					[{$\varphi = \delta_1 > 1$}, align=center, base=bottom, edge label={node [midway,above] {$\frac{2}{3}$} }
					[{$\varphi = (\delta_1 > 1) \wedge (\delta_1 < \delta_2)$}, align=center, base=bottom, edge label={node [midway,above] {$\frac{1}{6}$} }
					[{\texttt{True}}]]
					[{$\varphi = (\delta_1 > 1) \wedge \neg(\delta_1 < \delta_2)$}, align=center, base=bottom, edge label={node [midway,above] {$\frac{5}{6}$} }
					[{\texttt{False}}]]]
					[{$\varphi = \neg(\delta_1 > 1)$}, align=center, base=bottom, edge label={node [midway,above] {$\frac{1}{3}$} }
					[{\texttt{False}}]]]
				\end{forest}
			}
			\caption{Execution Tree}
			\label{fig:exam_tree}

		\end{subfigure}
		\caption{An example program and its symbolic execution tree}
		\label{fig:example}
	\end{figure}

	\begin{paragraph}{Example}
		Consider the code in Fig.~\ref{fig:exam_program} and suppose you wish to calculate the probability of the program returning \texttt{True}.
		%
		Following Alg.~\ref{alg:sample} for lines 1,2, we generate fresh probabilistic symbolic variables for $\mathtt{x}$ and $\mathtt{y}$, $\delta_1$ and $\delta_2$, respectively.
		%
		We also store the distributions which $\delta_1$ and $\delta_2$ are samples from, namely the discrete uniform distribution $\mathcal{U}\{1,3\}$.
		%
		In our notation, we say that $\sigma = \{ \mathtt{x} \mapsto \delta_1, \mathtt{y} \mapsto \delta_2 \}$ and $P = \{ \delta_1,\delta_2 \mapsto \mathcal{U}\{1,3\} \}$.
		%
		Now following Alg.~\ref{alg:branch} to process line 4 of Fig.~\ref{fig:exam_program}, note that $\mathcal{D} = \{1,2,3\} \times \{1,2,3\}$ and

		\begin{align*}
			p_c = \Pr[\delta_1 < \delta_2 \mid \delta_1 > 1] &= \frac{\Pr[(\delta_1 < \delta_2) \wedge (\delta_1 > 1)]}{\Pr[\delta_1 > 1]}\\
			&= \frac{\displaystyle\sum_{(v_1,v_2) \in \mathcal{D}} [(\delta_1 < \delta_2) \wedge (\delta_1 > 1)\{\delta_1 \mapsto v_1, \delta_2 \mapsto v_2\}]}{\displaystyle\sum_{(v_1,v_2) \in \mathcal{D}} [(\delta_1 > 1)\{\delta_1 \mapsto v_1\}]}\\
			&= \frac{1}{6}
		\end{align*}
	\end{paragraph}

	\subsection{Query Generation}
	\label{sec:query_gen}

	At this point, we have an execution tree where each branch is annotated with the probability of taking the true and false branch.
	%
	From this tree we construct a set of paths which are represented by a path condition and set of probabilities.
	%
	If there are $n$ paths, and $\varphi_i$ is the $i^\text{th}$ path condition, and $p_{i_j}$ is the probability of the $j^{\text{th}}$ branch in the $i^{\text{th}}$ path, then we have $\{ (\varphi_1, \{p_{1_1},\ldots,p_{1_{k_1}}\}), \ldots, (\varphi_n, \{p_{n_1},\ldots,p_{n_{k_n}}\})\}$.
	%
	To calculate the probability of an entire path, or equivalently, the probability of a path condition, $\varphi$, being true, we can simply multiple each of the branch probabilities.
	%
	Recall that $p_{i_j}$ is the \textit{conditional} probability of the $j^{\text{th}}$ branch condition being true along path $i$.
	%
	Since $\varphi_i$ is the conjunction of each of the $k_i$ branch conditions, the probability of taking the entire path can be computed using the rule $\Pr[A \wedge B] = \Pr[A \mid B]\Pr[B]$ where $A$ and $B$ are any two events.
	%
	In other words, since $\varphi_i = c_{i_1} \wedge \cdots \wedge c_{i_{k_i}}$, then
	\begin{align*}
		\Pr[\varphi_i] &= \Pr[c_{i_1} \wedge \cdots \wedge c_{i_{k_i}}]\\
		&= \Pr[c_{i_1} (c_{i_2} \wedge \cdots \wedge c_{i_{{k_i}-1}} \wedge c_{i_{k_i}})]\\
		&= p_{i_1} \cdot \Pr[c_{i_2} \wedge \cdots \wedge c_{i_{k_i}}]\\
		&\vdots\\
		&= \displaystyle\prod_{j=1}^{k_i} p_{i_j}
	\end{align*}

	So, if we let $p_i = \prod_{j=1}^{k_i} p_{i_j}$, then we can simplify our set of paths to $\Phi = \{ (\varphi_1, p_1), \ldots, (\varphi_n, p_n)\}$.
